Apache Flink\texttrademark: Unified Stream and Batch Processing \\ in a Single Engine

### 2015-12-02 batch.tex
Sent to Stephan, Kostas.

\subsection{Batch Iterations on top of Dataflows}
\label{sec:batch-iterations}
Structured iteration support is the basis of most distributed batch processing applications. Iterative graph analytics, parallel gradient descent and optimisation techniques have been implemented in the past on top of Bulk Synchronous Parallel (BSP) and Stale Synchronous Parallel (SSP) models, among others. The Flink execution model, allows for any type of structured iteration logic using iteration control events, as explained in \autoref{sec:iterations}. For example in the case of a BSP execution, iteration control events mark the beginning and the end of supersteps in an iterative computation. Finally, Flink introduces further novel optimisation techniques such as the concept of \emph{delta} iterations \cite{DBLP:journals/pvldb/EwenTKM12}, which can exploit sparse computational dependencies (already part of Gelly, Flink's Graph API).

>>>>

\subsection{Batch Iterations on top of Dataflows}
\label{sec:batch-iterations}
Iterative graph analytics, parallel gradient descent and optimisation techniques have been implemented in the past on top of Bulk Synchronous Parallel (BSP) and Stale Synchronous Parallel (SSP) models, among others. The Flink execution model, allows for any type of structured iteration logic to be implemented on top, by using iteration control events. For instance, in the case of a BSP execution, iteration control events mark the beginning and the end of supersteps in an iterative computation. Finally, Flink introduces further novel optimisation techniques such as the concept of \emph{delta} iterations \cite{DBLP:journals/pvldb/EwenTKM12}, which can exploit sparse computational dependencies (already part of Gelly, Flink's Graph API).

### 2015-12-02 exec-model.tex
Updates from ShareLaTeX
Sent to Stephan, Kostas.

>>>>

\subsection{Iterative Dataflows}
\label{sec:iterations}

\begin{figure}[t!]
  \centering
    \includegraphics[width=.75\textwidth]{figs/iterations.pdf}
    \vspace{-6mm}
  \caption{The iteration model of Apache Flink}
  \vspace{-2mm}
  \label{fig:iterations}
\end{figure}

Incremental processing and iterations are crucial for applications such as graph processing and machine learning. Support for iterations in data-parallel processing platforms typically relies on submitting a new job for each iteration or by adding additional nodes to a running DAG \cite{DBLP:journals/pvldb/BuHBE10, DBLP:conf/hotcloud/ZahariaCFSS10}, 
or feedback edges \cite{murray2013naiad}.

Iterations in Flink are implemented as  \emph{iteration steps}, special operators that themselves can contain an execution graph (\autoref{fig:iterations}). To maintain the (acyclic) DAG-based runtime and scheduler, Flink allows for iteration ``head'' and ``tail'' tasks that are \emph{implicitly} connected with feedback edges. The role of these tasks is to establish an active feedback channel to the iteration step and provide coordination for processing data records in transit within this feedback channel. Coordination is needed for implementing any type of structured parallel iteration model such as the Bulk Synchronous Parallel (BSP) model and is implemented using control event. We explain how iterations are implemented in the DataStream and DataSet APIs in \autoref{sec:stream-iterations} and  \autoref{sec:batch-iterations} respectively.

### 2015-12-02 streaming.tex
stream not streaming
Sent to Stephan, Kostas.

\subsection{Asynchronous Streaming Iterations}
Stream loops are essential for several applications. Common examples are feedback loops for incrementally building and training machine learning models, reinforcement learning and graph algorithms in the semi-streaming model \cite{feigenbaum2005graph}.
In most such cases, feedback loops need no coordination. Asynchronous iterations cover the communication needs for streaming applications and differ from parallel optimisation problems that are based on structured iterations on finite data. As presented, in \autoref{sec:iterations} and \autoref{fig:iterations} the execution model of Apache Flink already covers asynchronous iterations, when no iteration control mechanism is enabled. In addition, to comply with fault-tolerance guarantees feedback streams are treated as operator state within the implicit iteration head operator and are part of a global snapshot, as explained in detail in \cite{carbone2015lightweight}. The DataStream API allows for an explicit definition of feedback streams and can trivially subsume future support for structured loops over streams such as differential dataflows \cite{murray2013naiad}.

>>>>

\subsection{Asynchronous Stream Iterations}
\label{sec:stream-iterations}
Loops in streams are essential for several applications such as incrementally building and training machine learning models, reinforcement learning and graph algorithms in the semi-streaming model \cite{feigenbaum2005graph}.
In most such cases, feedback loops need no coordination. Asynchronous iterations cover the communication needs for streaming applications and differ from parallel optimisation problems that are based on structured iterations on finite data. As presented, in \autoref{sec:iterations} and \autoref{fig:iterations} the execution model of Apache Flink already covers asynchronous iterations, when no iteration control mechanism is enabled. In addition, to comply with fault-tolerance guarantees feedback streams are treated as operator state within the implicit iteration head operator and are part of a global snapshot, as explained in detail in \cite{carbone2015lightweight}. The DataStream API allows for an explicit definition of feedback streams and can trivially subsume future support for structured loops over streams such as differential dataflows \cite{murray2013naiad}.
